---
title: "Sequencing Run Report"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    mathjax: NULL
params:
    summary_df: 'overall.qc.csv'
    run_amplicons: FALSE
    version: ""
    revision: ""
    nf_version: ""
    neg_ctrl_substrings: ""
---

<!-- Style updates mostly to make sure the caption boxes are all matching -->
<style>
.navbar-brand {
  font-weight: bold;
}
.value-box .caption {
  font-size: 24px;
  color: white
}
.value-box .value {
    color: white;
}
</style>

```{r, setup, include=FALSE}
# Load in libraries
library(flexdashboard)
library(plotly)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(data.table)

# Set to the jpeg as its smaller
knitr::opts_chunk$set(dev = 'jpeg')
```

```{r, global, include=FALSE, echo=FALSE, error=TRUE}
# Create global variables and functions
# Set Table number to account for optional tables
table_number <- 0

# Read QC csv file
summary_df <- read.table(params$summary_df, header = TRUE, sep = ",")

# Specify and create a df with negative controls
control_substrings = trimws(strsplit(params$neg_ctrl_substrings, "[,;|]+")[[1]])
control_regex = paste(control_substrings, collapse = "|")
is_control <- grepl(
  control_regex,
  summary_df$sample,
  ignore.case = TRUE
)
has_controls <- any(is_control)
control_df <- summary_df[is_control, ]

# Genome completeness colour palette
get_genome_completeness_color <- function(value) {
  # Create the Red to Yellow to Green color palette for 100 values
  red_to_green_palette <- colorRampPalette(c("#FF0039", "#FF7518", "#3FB618"))(100)

  # Calculate the index of the color in the palette based on the input value
  if (value == 0) {
    value <- value + 0.00001
  }
  color_index <- ceiling(value * 100)

  # Return the corresponding color with alpha 0.7
  selected_color <- red_to_green_palette[color_index]
  selected_color <- adjustcolor(selected_color, alpha = 0.7)
  return(selected_color)
}

# Horizontal Line function for plotly
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash = "dash")
  )
}

# Create a new column based on the qc_pass value for colours
add_pass_value_int <- function(df, column) {
  # Split the column values on semicolons
  split_values <- strsplit(df[[column]], ';')

  # Define which int values go with which statement
  key_phrases <- c('PASS', 'POTENTIAL_FRAMESHIFTS', 'PARTIAL_GENOME')
  values <- c(1, 2, 3)

  # Initialize a vector to store the results
  results <- integer(length(df[[column]]))

  for (i in 1:length(split_values)) {
    # Check if any key phrase is contained in the split values
    matching_phrases <- intersect(split_values[[i]], key_phrases)

    if (length(matching_phrases) > 0) {
      # If matching phrases are found, assign the corresponding integer value
      results[i] <- min(values[match(matching_phrases, key_phrases)])
    } else {
      results[i] <- 0  # If none of the key phrases are found, assign 0
    }
  }

  # Return results
  return(results)
}
```

Run Summary
==============================

Row
------------------------------

### Total Samples
```{r, samples_box, echo=FALSE, error=TRUE}
total_samples <- nrow(summary_df)
valueBox(total_samples, icon = "")
```

### Total Passing Samples
```{r, passing_box, echo=FALSE, error=TRUE}
passing_samples <- sum(summary_df[["qc_pass"]] == "PASS", na.rm = FALSE)
valueBox(passing_samples, icon = "")
```

### Run Name
```{r, run_name_box, echo=FALSE, error=TRUE}
if ("run_name" %in% colnames(summary_df)) {
  run_name <- summary_df[1, 'run_name']
} else {
  run_name <- "NA"
}
valueBox(run_name, icon = "")
```

### Negative Control Status
```{r, controls_status_box, echo=FALSE, error=TRUE}
# Based on run satus, set colour and icon
run_status = summary_df[1, 'run_status']
if ( run_status == 'PASS' ) {
    sec_icon = ""
    sec_color = "success"
} else if ( run_status == 'WARN' ) {
    sec_icon = ""
    sec_color = "warning"
} else {
    sec_icon = ""
    sec_color = "danger"
}
valueBox(
  run_status,
  icon = sec_icon,
  color = sec_color
)
```

Row {dataset-height=250}
------------------------------

### {-}
```{r, plot_completeness, echo=FALSE, error=TRUE}
# All sample genome completeness plot
df <- select(summary_df, sample, genome_completeness)

# Creat plot, adding in titles and the PASS/WARN hlines (may need to figure out how to label these)
p <- plot_ly(df, x = ~sample, y=~genome_completeness, type = 'bar', color = ~genome_completeness, colors = "RdYlGn", name="") %>%
  layout(
    title = "Sample Genome Completeness",
    xaxis = list(title = "Sample Name"),
    yaxis = list (title = "Genome Completeness", tickformat = '.2%'),
    shapes = list(hline(0.9), hline(0.5))
  )
p
```

> **Figure 1.** Per Sample Genome Completeness. Comparison of the percentage of bases called in the consensus sequence of each sample.
> The dashed lines at 50% and 90% show the cut-offs for an incomplete genome and a partial genome respectively.

Row
------------------------------

### **Table `r table_number <- table_number + 1; as.roman(table_number)`.** All Samples Data Summary
```{r, data_table, echo=FALSE, error=TRUE}
# Create an interactive table of all samples with the different qc statuses highlighted
df <- summary_df %>%
  select(any_of(c('sample', 'qc_pass', 'mean_sequencing_depth', 'median_sequencing_depth',
                  'num_aligned_reads', 'total_variants', 'num_snps', 'num_iupacs',
                  'num_deletions', 'num_deletion_sites', 'num_insertions',
                  'num_insertion_sites', 'genome_completeness')))

# Rename column names
df <- df %>%
  rename(
    "Sample" = sample,
    "Mean Depth" = mean_sequencing_depth,
    "Median Depth" = median_sequencing_depth,
    "Total Aligned Reads" = num_aligned_reads,
    "Variant Count" = total_variants,
    "# SNPs" = num_snps,
    "# Del" = num_deletions,
    "# Del Sites" = num_deletion_sites,
    "# Ins" = num_insertions,
    "# Ins Sites" = num_insertion_sites,
    "Genome Completeness" = genome_completeness,
    "QC Pass" = qc_pass
  )

# Separate as its not an ONT output
if ("num_iupacs" %in% colnames(df)) {
  df <- df %>%
    rename(
      "# IUPACs" = num_iupacs
    )
}

# Add color to table
df$completeness_int_color <- add_pass_value_int(df, 'QC Pass')

# Using datatable, removing the index row and adding in highlights for the status
summary_dt <- df %>%
  DT::datatable(rownames = FALSE,
    options = list(
      scrollX = FALSE,
      columnDefs = list(
        list(className = 'dt-left', targets = "_all"),
        list(visible=FALSE, targets='completeness_int_color')
    ))) %>%
    formatStyle('completeness_int_color',
      target = 'row',
      backgroundColor = styleEqual(c(1, 2, 3, 0),
      c('', 'orange', '#FFFF66', '#FF6666')
    )) %>%
    formatStyle('Genome Completeness',
      background = styleColorBar(range(0,1), 'lightblue', angle = -90),
      backgroundSize = '88% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'left'
  )

summary_dt
```

Row
------------------------------

### {-}
```{r median-depth-plot, error=TRUE, echo=FALSE}
# Getting all the depth TSV files
files <- list.files("base_coverages", pattern = "\\.bed$", full.names = TRUE)

# Combine all dfs
long_df <- rbindlist(
  lapply(files, fread,
    select = c("reference_name", "position", "depth"),
    colClasses = c(
      reference_name = "character",
      position = "integer",
      depth = "integer"
    )
  )
)

# Summarize positions
#  Ymin and Ymax for drawing the stdev region
#  Ymin should not be below 0
summary_depth_df <- long_df %>%
  group_by(position) %>%
  summarise(
    mean = mean(depth, na.rm = TRUE),
    sd = sd(depth, na.rm = TRUE),
    median = median(depth, na.rm = TRUE),
    ymin = pmax(mean - sd, 0),
    ymax = mean + sd,
    .groups = "drop"
  )

# Get maximum genome_length for x-axis
genome_length <- max(summary_depth_df$position, na.rm = TRUE)

# Plot figure
plot_ly(summary_depth_df, x = ~position, text = ~paste0(
  "<b>Position: </b>", position,
  "<br><b>Mean ± SD: </b>", round(mean, 2), " ± ", round(sd, 2)
), hoverinfo = "text") %>%
  add_ribbons(ymin = ~ymin, ymax = ~ymax, fillcolor = "rgba(0, 0, 255, 0.25)", line = list(color = "transparent"), name = "Mean ± SD", hoverinfo = "skip") %>%
  add_lines(y =~median, line = list(color = "blue"), name = "Median read depth") %>%
  layout(title = list(text= "Median genomic read depth with deviation", x = 0.5),
    xaxis = list(title = "Genome Position", range = c(0, genome_length), nticks = 10, showgrid = FALSE, gridcolor = "#dadce0", griddash = "dot", tickformat = ".0f"),
    yaxis = list(title = "Median read depth", rangemode = "tozero", showgrid = FALSE),
    showlegend = FALSE
  )
```


> **Figure 2.** Median summarized genomic read depth with standard deviation. The median depth and standard deviation were calculated for each genomic position and plotted such that the median depth is highlighted plus and minus 1 standard deviation.

<!-- ---------------------------------------------------------- -->
<!-- Create Optional pages with child documents -->
<!-- ---------------------------------------------------------- -->

<!-- ------------------ -->
<!-- Negative Controls if given -->
<!-- ------------------ -->
```{r, neg_controls, child="sampleNegative.Rmd", params=params, eval=has_controls}
```


<!-- ------------------ -->
<!-- Amplicons if given -->
<!-- ------------------ -->
```{r, amplicons, child="sampleAmplicons.Rmd", params=params, eval=params$run_amplicons}
```

<!-- ---------------------- -->
<!-- Back to normal Rmd doc -->
<!-- ---------------------- -->

Tool Versions
==============================

## Row

### {-}

```{r pipeline-name}
valueBox('phac-nml/viralassembly', "Pipeline Name")
```

### {-}

```{r pipeline-version}
valueBox(params$version, "Pipeline Version")
```

### {-}

```{r pipeline-rev}
valueBox(params$revision, "Pipeline Revision")
```

### {-}

```{r nfversion}
valueBox(params$nf_version, "Nextflow Version")
```

## Row

### **Table `r table_number <- table_number + 1; as.roman(table_number)`.** Software Versions {data-height=550}

```{r, software_ver, echo=FALSE, error=TRUE}
# Read the YAML file
yaml_data <- yaml::read_yaml("software_versions.yml")

# Track to build df
process <- c()
softwares <- c()
versions <- c()

# Loop over the data to build df
for (key_process in names(yaml_data)) {
  process_hash <- yaml_data[[key_process]]

  for (software_name in names(process_hash)) {
    software_version <- process_hash[[software_name]]

    # Add to lists
    process <- c(process, key_process)
    softwares <- c(softwares, software_name)
    versions <- c(versions, software_version)
  }
}

# Create table
df <- data.frame("Process Name" = process, "Software" = softwares, "Version" = versions)
ddt <- df %>%
  DT::datatable(rownames = FALSE, class = 'cell-border stripe',
    options = list(
      scrollX = FALSE,
      columnDefs = list(
        list(className = 'dt-left', targets = "_all")
  )))
ddt
```
