---
title: "Sample Report"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    mathjax: NULL
params:
    sample_df: NULL
    min_allele_freq: 0.65
---

<!-- Style updates mostly to make sure the caption boxes are all matching -->
<style>
.value-box .caption {
  font-size: 24px;
  color: white
}
.value-box .value {
    color: white;
}
</style>

```{r, setup, include=FALSE}
# Load in libraries
library(flexdashboard)
library(plotly)
library(dplyr)
library(DT)
library(knitr)
library(ggplot2)
library(data.table)

# Set to the jpeg as its smaller
knitr::opts_chunk$set(dev = 'jpeg')

# Set parameters
sample_df <- fread(params$sample_df, header = TRUE)
sample <- sample_df$sample[[1]]
min_allele_freq <- params$min_allele_freq

# Horizontal Line function for plotly
hline <- function(y = 0, color = "black") {
  list(
    type = "line",
    x0 = 0,
    x1 = 1,
    xref = "paper",
    y0 = y,
    y1 = y,
    line = list(color = color, dash = "dash")
  )
}

# Genome completeness colour palette
get_genome_completeness_color <- function(value) {
  # Create the Red to Yellow to Green color palette for 100 values
  red_to_green_palette <- colorRampPalette(c("#FF0039", "#FF7518", "#3FB618"))(100)

  # Calculate the index of the color in the palette based on the input value
  if (value == 0) {
    value <- value + 0.00001
  }
  color_index <- ceiling(value * 100)

  # Return the corresponding color with alpha 0.7
  selected_color <- red_to_green_palette[color_index]
  selected_color <- adjustcolor(selected_color, alpha = 0.7)
  return(selected_color)
}
```

Row
------------------------------

### Sample Name
```{r, sample_name_box, echo=FALSE, error=TRUE}
valueBox(sample, icon = "")
```

### Genome Completeness
```{r, genome_completeness_box, echo=FALSE, error=TRUE}
completeness <- sample_df$genome_completeness[1]
sec_color <- get_genome_completeness_color(completeness)
valueBox(
    completeness,
    icon = "",
    color = sec_color
)
```

### Sample Status
```{r, sample_status_box, echo=FALSE, error=TRUE}
# Based on run satus, set colour and icon
sample_status <- sample_df$qc_pass[1]
if ( sample_status == 'PASS' ) {
    sec_icon = ""
    sec_color = "success"
} else if ( grepl(sample_status, 'POTENTIAL_FRAMESHIFTS') ) {
    sec_icon = ""
    sec_color = "orange"
} else if ( grepl(sample_status, 'PARTIAL_GENOME') ) {
    sec_icon = ""
    sec_color = "warning"
} else {
    sec_icon = ""
    sec_color = "danger"
}
valueBox(
  sample_status,
  icon = sec_icon,
  color = sec_color
)
```

## Row {data-height=170}

### **Table I:** Expanded Sample Summary Table
```{r, expanded_summary_table, echo=FALSE, error=TRUE}

# Select needed columns
df <- sample_df %>%
  select(any_of(c('sample', 'qc_pass', 'mean_sequencing_depth', 'median_sequencing_depth',
                  'num_aligned_reads', 'total_variants', 'num_consensus_n', 'num_snps', 'num_deletions',
                  'num_deletion_sites', 'num_insertions', 'num_insertion_sites',
                  'genome_completeness', 'minor_snps', 'minor_indels', 'nextclade_frameshifts', 'premature_stop_codons', 'mutated_stop_codons')))

# Rename columns
df <- df %>%
  rename(
    "Sample" = sample,
    "QC Pass" = qc_pass,
    "Mean Depth" = mean_sequencing_depth,
    "Median Depth" = median_sequencing_depth,
    "Total Aligned Reads" = num_aligned_reads,
    "Variant Count" = total_variants,
    "# Masked Sites" = num_consensus_n,
    "# SNPs" = num_snps,
    "# Del" = num_deletions,
    "# Del Sites" = num_deletion_sites,
    "# Ins" = num_insertions,
    "# Ins Sites" = num_insertion_sites,
    "Genome Completeness" = genome_completeness,
    "QC Pass" = qc_pass
  )

# Rename columns for minor snps, if available
if ("minor_snps" %in% colnames(df)) {
  df <- df %>%
    rename(
      "# Minor SNPs" = minor_snps,
      "# Minor Indels" = minor_indels
    )
}

# Rename columns for nextclade mutations, if available
if ("nextclade_frameshifts" %in% colnames(df)) {
  df <- df %>%
    rename(
      "Frameshifts" = nextclade_frameshifts,
      "Premature Stop Codons" = premature_stop_codons,
      "Mutated Stop Codons" = mutated_stop_codons
    )
}

# Create sample information table
sddt <- df %>%
  datatable(rownames = FALSE,
    options = list(
      dom = 't',
      paging = FALSE,
      scrollX = FALSE,
      scrollY = "",
      columnDefs = list(
        list(className = 'dt-left', targets = "_all")
      )
    )
  ) %>%
  formatStyle(
    'Genome Completeness',
    background = styleColorBar(range(0,1), 'lightblue', angle = -90),
    backgroundSize = '88% 88%',
    backgroundRepeat = 'no-repeat',
    backgroundPosition = 'left'
  )

sddt
```

## Row

### {-}
```{r, variant_plot, echo=FALSE, error=TRUE}
# Create table from bam file csv showing variants and subvariants
var_f <- paste0("./variation_csvs/", sample, "_variation.csv")
if (file.exists(var_f)) {
  variants <- read.table(var_f, header = TRUE, sep = ",")
} else {
    variants <- data.frame(matrix(ncol = 12, nrow = 1))
    colv <- c("position", "ref_base", "percentage_nonref", "variant_type", "total_reads", "A", "T", "C", "G", "del", "ins", "distribution")
    colnames(variants) <- colv
}

# Text info for hover
text_info <- paste(
  'Position: ', variants[['position']], '<br>',
  'Reference: ', variants[['ref_base']], '<br>',
  'Total Reads: ', variants[['total_reads']], '<br>',
  '  A: ', variants[['A']], '<br>',
  '  T: ', variants[['T']], '<br>',
  '  G: ', variants[['G']], '<br>',
  '  C: ', variants[['C']], '<br>',
  '  Del:', variants[['del']], '<br>',
  '  Ins:', variants[['ins']], '<br>',
  'Insertion Distribution: ', variants[['distribution']], '<br>',
  sep = ''
)
variants$text_info <- text_info

# Plot variation table
#  Also customdata instead of text due to https://github.com/plotly/plotly.R/issues/1859
p <- plot_ly(variants) %>%
  add_trace(x = variants[['position']], y = variants[['percentage_nonref']],
      type = 'scatter', mode = 'markers',
      color = ~factor(variant_type, levels = c("A SNP", "T SNP", "G SNP", "C SNP", "Ref", "Mixed", "Deletion", "Insertion")),
      marker = list(sizemode = 'diameter', size = 15.0, opacity = 0.7),
      customdata = ~text_info,
      hovertemplate = paste(
        '<b>Percentage</b>: %{y:.2f}<br>',
        '%{customdata}<extra></extra>',
        sep=''
      )
  )

# Add final layout to plot
p <- p %>%
  layout(
    xaxis = list (title = "Position", tickformat = ".0f"),
    yaxis = list (title = "Percent Non Reference Base", tickformat = '2.%'),
    shapes = list(hline(min_allele_freq * 100))
  )
p
```

> **Figure 1.** Variant Read Counts. The sample pileup was analyzed to find positions in the genome where
> the alternate allele percentage was greater-than or equal to 15%.

## Row

### **Table II:** Consensus Called Variants
```{r, variant_table, echo=FALSE, error=TRUE}
# Create table of called variants using the vcf file turned into a tsv
var_f <- paste0("./variant_tsvs/", sample, ".vcf.tsv")
if (file.exists(var_f)) {
  variants <- read.table(var_f, header = TRUE, sep = "\t", check.names = FALSE, colClasses = "character")
  ddt <- variants %>%
    DT::datatable(rownames = FALSE,
      options = list(
        scrollX = FALSE
    )) %>%
    formatRound(columns = c('Qual'), digits = 2) %>%
    formatRound(columns = c('Alt Allele Fraction'), digits = 4) %>%
    # Bars to show at a glance
    formatStyle(
      'Alt Allele Fraction',
      background = styleColorBar(range(0,1), 'lightblue', angle = -90),
      backgroundSize = '88% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'left'
    )
} else {
  # Create an empty table if file doesn't exist
  variants <- data.frame(matrix(ncol = 7, nrow = 0))
  colv <- c("Chrom", "Pos", "Ref", "Alt", "Qual", "Depth", "Alt Allele Fraction")
  colnames(variants) <- colv
  ddt <- variants %>%
    DT::datatable(rownames = FALSE)
}
ddt
```

Row
------------------------------
### {-}
```{r, depth_plot, echo=FALSE, error=TRUE}
cov_f <- paste0("./base_coverages/", sample, ".per_base_coverage.bed")
coverage <- read.table(cov_f, header = TRUE, sep = "\t")
coverage <- select(coverage, position, depth)
# Set depths of 0 to 1 to allow continuous visualization with log plot
coverage$depth[coverage$depth == 0] <- 1

# Create the background shapes needed
background_shapes <- list(
  list(
    type = "rect",
    fillcolor = "red",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 1,
    y1 = 10
  ),
  list(
    type = "rect",
    fillcolor = "yellow",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 10,
    y1 = 30
  ),
  list(
    type = "rect",
    fillcolor = "green",
    opacity = 0.2,
    x0 = min(coverage$position),
    x1 = max(coverage$position),
    y0 = 30,
    y1 = max(coverage$depth)
  )
)

# Create plot
depthp <- plot_ly(x = ~coverage$position, y = ~coverage$depth, type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
    type = "log",
    title = "Sequencing Depth (Log Scale)",
    tickvals = c(10, 30, 100, 1000),
    ticktext = c(10, 30, 100, 1000)
  ),
  xaxis = list(
    title = "Genomic Position (nt)",
    tickformat = ".0f"
  ),
  shapes = background_shapes
)

depthp
```
> **Figure 2.** Positional genomic depth. The Y-axis is log-scaled so positions with no genomic depth have been set to 1 to allow a continuous visualization. Zones have been marked using a colour scheme where: 0-10 is red corresponding to masked sites, 10-30 is yellow corresponding to sites where there is a basecall but limited depth, and 30+ is green for positions with acceptable or better depth.

## Row

### {-}
```{r, qual_plot, echo=FALSE, error=TRUE}
quality <- read.table(paste0("./base_qualities/", sample, ".baseq.stats.tsv"), sep = "\t", header = TRUE)
custom_tickvals <- c(0,10,20,30)

# Create the background shapes needed
background_shapes <- list(
  list(
    type = "rect",
    fillcolor = "red",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 0,
    y1 = 20
  ),
  list(
    type = "rect",
    fillcolor = "yellow",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 20,
    y1 = 30
  ),
  list(
    type = "rect",
    fillcolor = "green",
    opacity = 0.2,
    x0 = min(quality$pos),
    x1 = max(quality$pos),
    y0 = 30,
    y1 = max(quality$rms_baseq)
  )
)

# Plot
qualityp <- plot_ly(quality ,x = ~quality$'pos', y = ~quality$'rms_baseq', type = "scatter", mode = "lines", line = list(color = "black")) %>%
  layout(
    yaxis = list(
      tickvals = custom_tickvals,
      ticktext = c("0", "10", "20", "30"),
      title = "RMS Base Quality Score"),
    xaxis = list(
      title = "Genomic Position (nt)",
      tickformat = ".0f"
    ),
    shapes = background_shapes
  )

qualityp
```
> **Figure 3.** Average root-mean-square read base quality score per genomic position
