<!-- Dynamically add in negative control data if negative controls are used -->
Negative Controls
==============================

Row
------------------------------

### Total Negative Controls
```{r, total_controls, echo=FALSE, error=TRUE}
total_controls <- nrow(control_df)
valueBox(total_controls, icon = "")
```

### Negative Control Status
```{r, controls_status, echo=FALSE, error=TRUE}
# Based on run satus, set colour and icon
run_status = control_df[1, 'run_status']
if ( run_status == 'PASS' ) {
    sec_icon = ""
    sec_color = "success"
} else if ( run_status == 'WARN' ) {
    sec_icon = ""
    sec_color = "warning"
} else {
    sec_icon = ""
    sec_color = "danger"
}
valueBox(
  run_status,
  icon = sec_icon,
  color = sec_color
)
```

### Negative Control Info
```{r, controls_info, echo=FALSE, error=TRUE}
control_info <- control_df[1, 'run_summary']
sec_color <- "warning"
if ( is.na(control_info) ) {
  control_info <- "Meets Requirements"
  sec_color <- "success"
}
valueBox(
  control_info,
  color = sec_color,
  icon = ""
)
```

Column {.sidebar}
------------------------------
**Negative Control Information**

Negative controls are identified using the control identifiers specified by the user when running the pipeline using the --neg_ctrl_substrings parameter. For this sequencing run, the following identifiers were used to identify negative control samples: **`r paste(control_substrings)`**.

Negative controls are expected to contain little to no target sequence. A negative control is flagged if its genome completeness is ≥ 10%, indicating that a substantial portion of the reference genome was recovered.

The table here summarizes sequencing and QC metrics for all samples identified as negative controls. Unexpected sequencing signal in these samples may indicate contamination, non-specific amplification, or other run-related issues.

Future iterations of this pipeline aim to include more checks for negative controls.

Row {data-height=500}
------------------------------

### **Table 1.** Negative Controls
```{r, n_table, echo=FALSE, error=TRUE}
df <- control_df %>%
    select(any_of(c("sample", "neg_control_info", "mean_sequencing_depth", "median_sequencing_depth",
        "num_aligned_reads", "total_variants", "genome_completeness", "qc_pass"))
    ) %>%
    rename(
        "Sample" = sample,
        "Negative Control Info" = neg_control_info,
        "Mean Depth" = mean_sequencing_depth,
        "Median Depth" = median_sequencing_depth,
        "Total Aligned Reads" = num_aligned_reads,
        "Variant Count" = total_variants,
        "Genome Completeness" = genome_completeness,
        "QC Pass" = qc_pass
    )

# Add color to table
df$completeness_int_color <- add_pass_value_int(df, 'QC Pass')

# Using datatable, removing the index row and adding in highlights for the status
summary_dt <- df %>%
  DT::datatable(rownames = FALSE,
    options = list(
      scrollX = FALSE,
      columnDefs = list(
        list(className = 'dt-left', targets = "_all"),
        list(visible=FALSE, targets='completeness_int_color')
    ))) %>%
    formatStyle('completeness_int_color',
      target = 'row',
      backgroundColor = styleEqual(c(1, 2, 3, 0),
      c('', 'orange', '#FFFF66', '#FF6666')
    )) %>%
    formatStyle('Genome Completeness',
      background = styleColorBar(range(0,1), 'lightblue', angle = -90),
      backgroundSize = '88% 88%',
      backgroundRepeat = 'no-repeat',
      backgroundPosition = 'left'
  )

summary_dt
```
